Symbolic Dynamic Programming for Continuous State and Observation POMDPs

### 2012-06-01 trunk/cpomdp/camdp-xadd/src/doc/NIPS/draft4-10page.tex

At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step. The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)\mathcal{T}(xd'_s,xd_s,a)\mathcal{Z}(xd'_s,xd_o,a)$. \footnote{For simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$)}
The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
{\footnotesize
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
\end{equation}
}
Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
%The decision at each time step is to pick an action based on past actions and observations. 
The possible set of conditional plans $\beta^h$ correspond to a decision tree. Figure 1 left, shows the decision tree for the simple case of three horizons with discrete observations. It represents a conditional plan consisting of an action and two possible observation strategies. For continuous observations, conditional plans are decision trees with infinitely many branches. In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1.%From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma^h :=\emptyset$\;
       \ForEach {$b_i \in B$}
       {
       	\ForEach {$a \in A$}
      	 {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		\If {$ContObs =true$}
       			{$(O^h,P(xd_{o_i}|xd_{s_i})) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'},:=\,$ \remapWithPrimes{$\alpha_j^{h}$} 
   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(xd_{o_i}|xd_{s_i}) \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
       	 }  
        	%monahan's pruning first generates all vectors, then prunes
              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
      }
             % $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
             % $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;
      \ForEach {$b_i \in B$}
      {
     	$\alpha_{b_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \Gamma^{h}$\;
     	$\Gamma_{PBVI}^h :=\ \Gamma^h \cup \alpha_{b_i}^h$\;
      }

@@

Calculating all possible ways of building the value function independent of a certain belief state can also define the optimal value function. To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}. Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as 
$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
{\footnotesize
\begin{align}
g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \\
\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
\Gamma^h  &= \bigcup_a \Gamma^h_a 
\end{align}
}
where $r_a = r(x_s,d_s,a)$.  
%Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $. The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.  
Lines 7--15 of the \textsc{PBVI} algorithm show the general algorithm used for both discrete and continuous observations. We will discuss this according to our symbolic framework in the next section. 
For all $\alpha$-vectors in the current $\Gamma$-set the \textsc{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}. A pairwise dominance test is also performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
%Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors.

@@

For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution represented in \textsc{PBVI}. The same approach is required for continuous observations only based on point-based dynamic programming backups.

Given an initial set of beliefs, we can compute the relevant observation partitions explained in the next section and then continue in a similar fashion of having discrete observations. %To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs. Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons. 
Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2. Also a set of beliefs are given as the input as often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in B$.
For each iteration, lines 11--12 product the following equation: 
{\footnotesize
\vspace{-2mm}
\begin{align}
g^h_{a,o,j} =  \int_{x_{s'}} \oplus_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
\end{align}
}
where the first probability is from equation (2) and the next probability is from equation (1) of the example. The $\alpha$-vectors are case functions primed using substitution in \textsc{Prime}. \texttt{Backup} performs regression on the $\alpha$-vectors where \textit{Boolean regression} is performing a symbolic restriction and \textit{Continuous regression} is a symbolic substitution, both SDP operations. 
Lines 14--16 cover closed form operations on the $\Gamma$-set such as cross-sum and union. Finally lines 18--20 find the maximum $\alpha$-vector within the $\Gamma^h$ set and add this to the final $\Gamma_{PBVI}^h$ if it is not already in the set. 
%Line 13 takes the cross-sum of the two sets, followed by adding the reward to the discounted result. For all actions the  $\Gamma$-set is added to the $\Gamma^h$ of the current horizon. The result is pruned as explained earlier.%which takes every possible combination of observations according to the $\alpha$-vectors. 
%All vectors are then multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
%The same operation is performed for all actions and then in 
%Line 15 computes the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
%The $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon. The action from which this $\alpha$-vector was chosen from is the maximum policy. 
%This operation is performed for each of the belief points $b_i \in B$ and for each belief point one dominant $\alpha$-vector is stored in the $\Gamma$-set. The algorithm iterates until convergence. 
In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.

\subsection{PBVI for DC State and DC Observations} 
The main operations required for value iteration in DC-POMDP is generating relevant observations to perform the bellman backup operation. 
The \texttt{GenRelObs} algorithm defines the steps required to generate these partitions. 
We use the second iteration of the \texttt{Power Plant} example to demonstrate partitioning of the entire continuous observation space. The main assumption is to consider some belief state and obtaining the relevant partitions according to that belief. We choose these beliefs such that they cover the entire probable state space; e.g. $b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$.

Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$. It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon:
$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
This operation of line 3 in \texttt{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model. For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section. 
Assuming that after $h=1$ the set of $\alpha$-vectors are defined as:
{\footnotesize
\vspace{-3mm}
\begin{align}
\alpha_1^1(t_s) &= 
\begin{cases}
 (t_s>200) &: -1000 \\
\neg(t_s>200) &: 100 \\
\end{cases}
\hspace{10mm} 
\alpha_2^1(t_s) = \top: -1 \nonumber
\end{align}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{substitute}{Substitute}

\Begin
{
	$z:=0$;\	
		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
		{
    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) \cdot P(xd'_{s_i}| xd_{s_i},a)\cdot \alpha_j $\;
    	\emph{// Point-based backup operation}\\
		}  
		\For {all $j$ in $\alpha_j(xd_s,xd_o)$}    
		{
		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i \cdot \alpha_j(xd_s,xd_o)$\}\ \;
		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
		}
		\For {all $j$ in $\delta_{j}(xd_o)$}    
		{
		$z := max(\delta_j(xd_o),z)$\;
		\emph{// Finding partition points on the observation space}\\
    	}
    	\emph{// $\phi_{z_k}$ logical term on each partition of $z$}\\
   		$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
   		\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
    \Return{$P(z)$} \;
    %do this for each belief in B
}
\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(z)$ }
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For $h=2$ the resulting $\alpha$-vectors after this regression step are functions of the observation and current state: 
{\footnotesize
\vspace{-2mm}
\begin{align}
\alpha_1^2(t_s,t_o) &= 
\begin{cases}
 (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\hspace{2mm} 
\alpha_2^2(t_s,t_o) &= \begin{cases}
(t_s - 10 < t_o<t_s) &: -0.1 \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\nonumber
\end{align}
}
We need the $\alpha$-vectors to partition on the observation space, thus next we eliminate the current state variable in line 4--5. During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.  The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
For $b_1: U[t_s;2,6]$ assume the following simple $\delta$-functions:
%the two vectors above the resulting $\delta$-functions are defined below:
{\footnotesize
\vspace{-2mm}
\begin{align}
\delta_{1}(t_o) &= 
\begin{cases}
 (-\infty < t_o< \infty) &: -0.4*t_o +6.8\\
\end{cases}
\hspace{5mm} 
\delta_{2}(t_o) &= \begin{cases}
 (-\infty < t_o< \infty) &: 1.3*t_o +3.4\\
\end{cases}
\nonumber
\end{align}
\vspace{-5mm}
}
%{\footnotesize
%\begin{align}
%\delta_{1}(t_o) &= 
%\begin{cases}
% (2<t_o<6) &: 15 - 2.5*t_o \\
%(-4<t_o<2) &: 10 \\
%(-8<t_o<-4) &: 20 + 2.5*t_o 
%\end{cases}
%\hspace{5mm} 
%\delta_{2}(t_o) &= \begin{cases}
% (2<t_o<6) &: -0.15 + 0.025*t_o \\
%(-4<t_o<2) &: -0.1 \\
%(-8<t_o<-4) &: -0.2 - 0.025*t_o 
%\end{cases}
%\nonumber
%\end{align}
%}

>>>>

At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step. The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)\mathcal{T}(xd'_s,xd_s,a)\mathcal{Z}(xd'_s,xd_o,a)$. \footnote{For simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$)}
The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
{\footnotesize
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
\end{equation}
}
Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
%The decision at each time step is to pick an action based on past actions and observations. 
The possible set of conditional plans $\beta^h$ correspond to a decision tree. Figure 1 left, shows the decision tree for the simple case of three horizons with discrete observations. It represents a conditional plan consisting of an action and two possible observation strategies. For continuous observations, conditional plans are decision trees with infinitely many branches. In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1.%From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma_{PBVI}^h :=\emptyset$\;
       \ForEach {$b_i \in B$}
       {
       	\ForEach {$a \in A$}
      	 {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		\If {$ContObs =true$}
       			{$(O^h,P(xd_{o_i}|xd_{s_i})) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'},:=\,$ \remapWithPrimes{$\alpha_j^{h}$} 
   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(xd_{o_i}|xd_{s_i}) \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
       	 }  
        	%monahan's pruning first generates all vectors, then prunes
              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
      }
             % $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
             % $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;
      \ForEach {$b_i \in B$}
      {
     	$\alpha_{b_i}^h :=\ \argmax_{\alpha_j \in \Gamma^h} \Gamma^{h}$\;
     	$\Gamma_{PBVI}^h :=\ \Gamma_{PBVI}^h \cup \alpha_{b_i}^h$\;
      }

@@

Calculating all possible ways of building the value function independent of a certain belief state can also define the optimal value function. To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}. Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as 
$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
{\footnotesize
\begin{align}
g^h_{a,o,j} &=  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \\
\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
\Gamma^h  &= \bigcup_a \Gamma^h_a 
\end{align}
}
where $r_a = r(x_s,d_s,a)$.  
%Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $. The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.  
Lines 7--15 of the \texttt{PBVI} algorithm show the general algorithm used for both discrete and continuous observations. We will discuss this according to our symbolic framework in the next section. 
For all $\alpha$-vectors in the current $\Gamma$-set the \texttt{Prune} function uses an LP-solver to prune any inconsistency in the branches of this vector in line 16 similar to the approach in ~\cite{penberthy94}. A pairwise dominance test is also performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
%Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors.

@@

For discrete and continuous states and discrete observations, we use the symbolic version of Monahan's algorithm as the DC-POMDP solution represented in \texttt{PBVI}. The same approach is required for continuous observations only based on point-based dynamic programming backups.

Given an initial set of beliefs, we can compute the relevant observation partitions explained in the next section and then continue in a similar fashion of having discrete observations. %To visualize we explain of the value iteration algorithm as defined in \texttt{VI} for DC-POMDPs. Each step of the algorithm is explained using the \textsc{Power plant} example for the first three horizons. 
Note that if the input to this algorithm is continuous observations $ContObs=true$ then line 8 finds the relevant observation partitions, else discrete observations are assumed and their probabilities are given in the problem specifications such as section 2. Also a set of beliefs are given as the input as often several beliefs are required to cover the belief space and thus the algorithm is executed for each one of $b_i \in B$.
For each iteration, lines 11--12 product the following equation: 
{\footnotesize
\vspace{-2mm}
\begin{align}
g^h_{a,o,j} =  \int_{x_{s'}} \large \oplus_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s) \nonumber
\end{align}
}
where the first probability is from equation (2) and the next probability is from equation (1) of the example. The $\alpha$-vectors are case functions primed using substitution in \texttt{Prime}. \texttt{Backup} performs regression on the $\alpha$-vectors where \textit{Boolean regression} is performing a symbolic restriction and \textit{Continuous regression} is a symbolic substitution, both SDP operations. 
Lines 14--16 cover closed form operations on the $\Gamma$-set such as cross-sum and union. Finally lines 18--20 find the maximum $\alpha$-vector within the $\Gamma^h$ set and add this to the final $\Gamma_{PBVI}^h$ if it is not already in the set. 
%Line 13 takes the cross-sum of the two sets, followed by adding the reward to the discounted result. For all actions the  $\Gamma$-set is added to the $\Gamma^h$ of the current horizon. The result is pruned as explained earlier.%which takes every possible combination of observations according to the $\alpha$-vectors. 
%All vectors are then multiplied by the discount $\gamma$ and added with the reward in line 14 where the addition is an augmented $\oplus$ for adding each vector with the reward function. 
%The same operation is performed for all actions and then in 
%Line 15 computes the $\Gamma$ set for each action are added to the set of $\Gamma^h$ to build all the vectors for the next horizon. 
%The $\alpha$-vector that maximizes the value-function with respect to this belief and save it as the new $\Gamma$-set for the next horizon. The action from which this $\alpha$-vector was chosen from is the maximum policy. 
%This operation is performed for each of the belief points $b_i \in B$ and for each belief point one dominant $\alpha$-vector is stored in the $\Gamma$-set. The algorithm iterates until convergence. 
In the next section we discuss the main part of our algorithm which is generating relevant observation partitions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{substitute}{Substitute}

\Begin
{
	$z:=0$;\	
		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
		{
		\emph{// Point-based backup operation}\\
    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) \cdot P(xd'_{s_i}| xd_{s_i},a)\cdot \alpha_j $\;
		}  
		\For {all $j$ in $\alpha_j(xd_s,xd_o)$}    
		{
		\emph{// Generate value of $\alpha$-vector for belief $b_i$ as a function of observations}\\
		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i \cdot \alpha_j(xd_s,xd_o)$\}\ \;
		}
		\For {all $j$ in $\delta_{j}(xd_o)$}    
		{
		\emph{// Finding partition points on the observation space}\\
		$z := max(\delta_j(xd_o),z)$\;
    	}
    	\emph{// $\phi_{z_k}$ logical term on each partition of $z$}\\
    	\emph{// Generate probabilities and partitions of observations that distinguish optimal $\alpha$-vector (conditional plan)}\\
   		$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
    \Return{$P(z)$} \;
    %do this for each belief in B
}
\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(z)$ }
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{PBVI for DC State and DC Observations} 
The main operations required for value iteration in DC-POMDP is generating relevant observations to perform the bellman backup operation. 
The \texttt{GenRelObs} algorithm defines the steps required to generate these partitions. 
We use the second iteration of the \textsc{Power Plant} example to demonstrate partitioning of the entire continuous observation space. The main assumption is to consider some belief state and obtaining the relevant partitions according to that belief. We choose these beliefs such that they cover the entire probable state space; e.g. $b_1: U[t_s;2,6]$ and $b_2: U[t_s;6,11]$.

Starting with an empty set of relevant observation partitions $z$, the algorithm first eliminates the next state variable $xd'_s$. It multiplies the observation model and the transition model in the set of $\alpha$-vectors of the previous horizon:
$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $. 
This operation of line 3 in \texttt{GenRelObs} is similar to that of \texttt{Regress} multiplied by the observation regressed model. For continuous variables the continuous regression and for discrete variables boolean restriction is used as defined in the previous section. 
Assuming that after $h=1$ the set of $\alpha$-vectors are defined equal to the reward function of equation (1).
%{\footnotesize
%\vspace{-3mm}
%\begin{align}
%\alpha_1^1(t_s) &= 
%\begin{cases}
% (t_s>200) &: -1000 \\
%\neg(t_s>200) &: 100 \\
%\end{cases}
%\hspace{10mm} 
%\alpha_2^1(t_s) = \top: -1 \nonumber
%\end{align}
%}
For $h=2$ the resulting $\alpha$-vectors after the regression step are functions of the observation and current state: 
{\footnotesize
\vspace{-2mm}
\begin{align}
\alpha_1^2(t_s,t_o) &= 
\begin{cases}
 (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\hspace{2mm} 
\alpha_2^2(t_s,t_o) &= \begin{cases}
(t_s - 10 < t_o<t_s) &: -0.1 \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\nonumber
\end{align}
}
We need the $\alpha$-vectors to partition on the observation space, thus next we eliminate the current state variable in line 4--5. During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.  The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
For $b_1: U[t_s;2,6]$ assume the following simple $\delta$-functions:
%the two vectors above the resulting $\delta$-functions are defined below:
{\footnotesize
\vspace{-2mm}
\begin{align}
\delta_{1}(t_o) &= 
\begin{cases}
 (-\infty < t_o< \infty) &: -0.4*t_o +6.8\\
\end{cases}
\hspace{5mm} 
\delta_{2}(t_o) &= \begin{cases}
 (-\infty < t_o< \infty) &: 1.3*t_o +3.4\\
\end{cases}
\nonumber
\end{align}
\vspace{-5mm}
}
%{\footnotesize
%\begin{align}
%\delta_{1}(t_o) &= 
%\begin{cases}
% (2<t_o<6) &: 15 - 2.5*t_o \\
%(-4<t_o<2) &: 10 \\
%(-8<t_o<-4) &: 20 + 2.5*t_o 
%\end{cases}
%\hspace{5mm} 
%\delta_{2}(t_o) &= \begin{cases}
% (2<t_o<6) &: -0.15 + 0.025*t_o \\
%(-4<t_o<2) &: -0.1 \\
%(-8<t_o<-4) &: -0.2 - 0.025*t_o 
%\end{cases}
%\nonumber
%\end{align}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\vspace{-6mm}
\centering
\hspace{-17mm}
\includegraphics[width=0.41\textwidth]{pics/beliefs.pdf}
\hspace{-12mm}
\includegraphics[width=0.42\textwidth]{pics/delta1.pdf}
\hspace{-12mm}
\includegraphics[width=0.42\textwidth]{pics/delta2.pdf}
\hspace{-17mm}
\vspace{-10mm}
\caption{\footnotesize 
{\it (left)} Beliefs $b_1,b_2$ for 1D \textsc{Power Plant} example; 
{\it (middle)} Observation dependent function for $b_1$ that partition the observation space into 2 regions with different probabilities $P(z_i)$ ; 
{\it (right)} 3 Relevant observation partitions and their probabilities for $b_2$.
}
\label{fig:timeSpace}
\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$. According to continuous observation space defined in \cite{pascal_ijcai05}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space. In their work, numerical solutions are proposed to find the roots of any two observation dependent function. Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximum in line 6--7: 
{\footnotesize
\vspace{-2mm}
\begin{align}
\mathrm{max} \Bigg(\delta_{1}(t_o),\delta_{2}(t_o)\Bigg) &= 
\begin{cases}
 z_1:(t_o<2 : -0.4*t_o +6.8) \\
z_2:(t_o>2 : 1.3*t_o +3.4)
\end{cases}
\nonumber
\end{align}
}
The two partitions define observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the two distinct observation partitions is found. If we integrate out the state and observation from the observation model and the belief state regardless of a particular $\alpha$-vector, it integrates to one: 
$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$
Thus we only need to multiply the indicators of each case partition as as in line 9 in this formula to obtain the probability of each partition such that it sums to one: 
$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
For our example the probabilities are defined as $P(z_1)=0.2$ and $P(z_2)=0.8$.
%{\footnotesize
%\begin{align}
%P(z_k)=
%\begin{cases}
% (t_o<2) &: 0.2 \\
%(2<t_o) &: 0.8\\
%\end{cases} 
%\nonumber
%\end{align}
%}

Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in Monahan's algorithm. 
%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
Next we present some results for 2-Dimensional continuous observation spaces.
